An Investigation of the SPDY Protocol

### 2012-12-19 paper/paper.tex
Update section reference in paper

\title{An Analysis of the SPDY Protocol as a Candidate for HTTP/2.0}
\date{\today}
\author{Brian Stack\thanks{ bis12@case.edu, Case Western Reserve University,
Cleveland, OH 44106}, Tom Dooner\thanks{ ted27@case.edu }}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SPDY}
\label{sec:background/spdy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The most interesting and likely to succeed proposal, SPDY is rather mature
at this point in time, having been implemented in both the most recent Chrome
and Firefox browsers in addition to having modules and patches for both the
Apache and Nginx web servers which allow them to speak SPDY.

@@

\begin{description}
\item[Framing Layer] This binary protocol provides a layer on top of a reliable transport
protocol (recommended to be TCP) and provides a way for HTTP request/response
pairs to be issued and received out of order over a single transport layer
connection.  This provides an expected performance improvement of HTTP/1.1 by
not requiring multiple connections to the host to be made in order to
parallelize resource acquisition. There are 2 types of frames, first is control
frames which are responsible for managing the lifecycle of a stream, which we
will define soon.  Next is the data frame which is used simply to transmit data
also allowing for streaming.
\item[Streaming] Streams are bidirectional and are used to send out of order
responses to requests. Importantly, streams can have priorities set by the
creator of the stream.  This, in conjunction with server-push (defined next)
allows for more important assets (images, javascripts, stylesheets, etc.) to be
sent over the wires first.  An example of the usefulness of this feature is a
webpage with many images that are unimportant for the actual content and layout
of the page.  In that case, the images could be given a low priority and sent
last.
\item[Server-Push] Once a server receives a request for an asset, if it is aware
of closely connected assets, it may send them behind the original response,
without waiting for the client to request them.  This will pre-populate the
cache of the client and save a round trip time for each item.
\item[Header Compression with Dictionary] If there are many small assets on a
webpage, then a significant portion of the transmitted bytes will be due to HTTP
headers rather than the actual content.  If this can be compressed, much
bandwidth is saved~\cite{binoy}.
\end{description}

% TODO: Expand on these for the full paper!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Speed+Mobility}
\label{sec:background/s+m}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This proposal~\cite{sm} comes from Microsoft.  It shares much in common with
SPDY, but with a few major differences. First is that rather than defining
their own session layer, Microsoft S+M relies on WebSockets for a session
layer.  In addition (this is very important), S+M uses HTTP to negotiate the
protocol upgrade rather than SPDY's use of SSL/TLS to do this same thing. This
adds quite a bit of latency to beginning the actual communication and is one of
the reasons that Facebook endorsed SPDY~\cite{fbook}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Web Asset Scraper}
\label{sec:research/scraper}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The server described in \S\ref{sec:research/server} needs to correctly simulate
real world webpages. In order to achieve this, we created a web scraping tool
which scraped websites chosen from the Alexa most popular websites
list\footnote{\url{http://www.alexa.com}} and recorded the size and number of
assets into a database.  This database will be accessed by the server upon each
request in order to determine which data to return.  In order to accurately
represent the original websites (and the state of the web), we employed a
headless web browser PhantomJS\footnote{\url{http://phantomjs.org}} which
mimicked a user's browser and thus requested all assets needed to render a web
page, even the dynamically loaded ones buried in CSS rules or Javascript.

We desire to record information that will allow us to assess the performance
characteristics of SPDY for each website we scrape. We recorded, for each
asset, the source IP we downloaded the asset from, the SSL certificate (if any)
for the connection on which we downloaded the asset, and the asset's filename
and filesize.

Not every website supports SSL, and the Alexa list ranks websites by hostname
combining SSL and non-SSL traffic. Since we are not measuring the time to
download the asset, we can default to SSL for websites that support it, as we
can determine useful information about the owner of assets from the SSL
certificate. For sites that do not support SSL, we will request the non-SSL
HTTP hostname.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Asset Characterization}
\label{sec:assets}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SPDY Combination Strategy}
SPDY's most complex new feature is the ability to multiplex requests for assets
over a single TCP connection. In this way, a web page's assets can benefit from
a warmed-up TCP connection which has passed the beginning of the TCP slow start
algorithm. Since SPDY is currently built atop SSL, SPDY can only combine assets
with a mutually-agreeable SSL certificate. The assets will be combined into
separate, interleaved streams over the same TCP connection, thus ideally
obtaining complete utilization of the connection and the benefit from TCP's
eventual discovery of an accurate congestion window.

In order to measure the efficiency of SPDY's asset combination, we must
determine what metric benefits from combining assets. The authors of SPDY claim
that the main benefit from asset combination is the reduction of network
connections required~\cite{spdy-whitepaper}.

Although the HTTP/1.1 specification allows for up to two simultaneous
connections to servers, modern browsers willfully violate it in service of page
load speed and will happily open up to six connections to each host. Websites
have also been pushing the limits by creating semantically equivalent hostnames
for asset servers (i.e. assets1.example.com, assets2.example.com, etc.) so
browsers will open six connections to each. Thus, browsers may wind up opening
hundreds of connections immediately upon loading a page -- thus dishonoring the
intent of TCP slow start.

SPDY's asset combination addresses this. In this section we will measure the
difference in the number of connections required to load each page under two
different assumptions: first that site owners could combine all assets with
mutually-agreeable SSL certificates into one SPDY connection and second that
site owners would combine all assets served by the same IP address \textit{and}
with mutually-agreeable SSL certificates into one SPDY connection. Logically,
the first method is an upper bound for the number of connections we can save,
we estimate that the second method is more accurate for most websites but is a
lower bound for websites which employ CDNs or asset host sharding.

\subsection{SSL-only Asset Combination}
\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth,angle=270]{plots/asset_combination_only_ssl_ratio.eps}
\caption{Connection Ratio -- SSL-only Combination Strategy}
\label{fig:combination-ssl}
\end{figure}
The results in Figure \ref{fig:combination-ssl} assume that all assets
described by a mutually-compatible SSL certificate are combined into one SPDY
connection. Two assets have \textit{mutually-compatible} SSL certificates if
their SSL certificates are valid for the host of the other asset (with either
the Canonical Name or the Alternative Hostnames). In this way, we
systematically combine assets by adding them to the largest group with which
they are compatible. We then compare the number of asset groups (SPDY
connections) with the number of assets (HTTP connections). The CDF of the ratio
for the Alexa top 100,000 websites is plotted.

As mentioned previously, this is an upper-bound for multiplexing savings.
Results nonetheless are dramatic: about 90\% of sites could load using
one-fifth the connections if combining assets via SPDY.

\subsection{SSL-and-IP Asset Combination}
\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth,angle=270]{plots/asset_combination_ssl_and_ip_ratio.eps}
\caption{Connection Ratio -- SSL-and-IP Combination Strategy}
\label{fig:combination-ssl-and-ip}
\end{figure}
The results in Figure \ref{fig:combination-ssl-and-ip} assume that all assets
described by a mutually-compatible SSL certificate and that are downloaded from
the same IP in practice are combined into one SPDY connection.

Figure \ref{fig:combination-ssl-and-ip} shows that, in the common-to-worst
case, 90\% of sites will halve the number of connections by switching to SPDY
and 40\% could load using one-fifth the connections.

\subsection{Connection Sizes}
\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth]{plots/asset_combination_connection_sizes.eps}
\caption{Connection Sizes After Asset Combination}
\label{fig:connection-sizes}
\end{figure}

Figure \ref{fig:connection-sizes} describes the size of connections when all
assets are combined together as described in the SSL-and-IP combination
strategy. 70\% of SPDY connections are noticeably larger than HTTP connections,
up to almost an order of magnitude in some cases. Some assets do not benefit
from SPDY's combination: small assets like Google Analytics tracking
Javascript, or external libraries used on web pages such as jQuery. These
comprise the left part of the SPDY curve.

% we measure two approaches: splitting on SSL cert and IP and simply splitting
% on SSL cert (as an upper bound for combination). The first method, ssl + ip,
% is prone to some error due to the use of CDNs confusing our scraping of
% assets.

>>>>

\title{An Investigation of the SPDY Protocol}
\date{\today}
\author{Brian Stack\thanks{ bis12@case.edu, Case Western Reserve University,
Cleveland, OH 44106}, Tom Dooner\thanks{ ted27@case.edu }}

@@

% TODO: Expand on these for the full paper!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SPDY}
\label{sec:background/spdy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The most interesting and likely to succeed proposal, SPDY is rather mature
at this point in time, having been implemented in both the most recent Chrome
and Firefox browsers in addition to having modules and patches for both the
Apache and Nginx web servers which allow them to speak SPDY.

@@

\begin{description}
\item[Framing Layer] This binary protocol provides a layer on top of a reliable transport
protocol (recommended to be TCP) and provides a way for HTTP request/response
pairs to be issued and received out of order over a single transport layer
connection.  This provides an expected performance improvement of HTTP/1.1 by
not requiring multiple connections to the host to be made in order to
parallelize resource acquisition. There are 2 types of frames, first is control
frames which are responsible for managing the lifecycle of a stream, which we
will define soon.  Next is the data frame which is used simply to transmit data
also allowing for streaming.
\item[Streaming] Streams are bidirectional and are used to send out of order
responses to requests. Importantly, streams can have priorities set by the
creator of the stream.  This, in conjunction with server-push (defined next)
allows for more important assets to be sent over the wires first.  An example of
the usefulness of this feature is a webpage with many images that are
unimportant for the actual content and layout of the page.  In that case, the
images could be given a low priority and sent last.
\item[Server-Push] Once a server receives a request for an asset, if it is aware
of closely connected assets, it may send them behind the original response,
without waiting for the client to request them.  This will pre-populate the
cache of the client and save a round trip time for each item.
\item[Header Compression with Dictionary] If there are many small assets on a
webpage, then a significant portion of the transmitted bytes will be due to HTTP
headers rather than the actual content.  If this can be compressed, much
bandwidth is saved~\cite{binoy}.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Speed+Mobility}
\label{sec:background/s+m}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This proposal~\cite{sm} comes from Microsoft.  It shares much in common with
SPDY, but with a few major differences. First is that rather than defining
their own session layer, Microsoft S+M relies on WebSockets for a session
layer.  In addition (this is very important), S+M uses HTTP to negotiate the
protocol upgrade rather than SPDY's use of SSL/TLS to do this same thing. This
adds quite a bit of latency to beginning the actual communication and is one of
the reasons that Facebook endorsed SPDY~\cite{fbook}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Web Asset Scraper}
\label{sec:research/scraper}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The server described in \S\ref{sec:research/server} needs to correctly simulate
real world webpages. In order to facilitate this, websites from the Alexa
rankings\footnote{\url{http://www.alexa.com}} have been scraped and the size and
number of assets on each recorded.  This database will be accessed by the server
upon each request in order to determine which data to return.  Great
consideration was taken to make sure that this data accurately represents the
state of the web, including executing javascript from each page in a headless
browser in order to find dynamically loaded assets. This is not part
of the proposed project as the work has already been completed in conjunction
with another student.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Asset Characterization}
\label{sec:assets}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
here is where we will discuss the asset measurement stuff.
